\thispagestyle{pagestyle}

\begin{center}
    \textbf{\fontsize{20pt}{24pt} \selectfont Abstract}
\end{center}

Developers are in certain need of a code editor that allows them to write their programs efficiently, without having to do any additional editor configuration or manage third-party dependencies. Small-scale projects do not require the use of feature-bloated integrated development environments, which can add complexity and make the entire coding process more difficult. A good integrated development environment needs to provide the bare minimum required for decent coding, and this can be pretty difficult to achieve nowadays, as programmers rely on more and more tools to enhance the development process. The variety of technologies and frameworks doesn't help either. Developers won't rely on a single development stack when building an application, so a code editor must always be ready to support multiple languages and be constantly improved and extended.

Building a general-purpose IDE that is also easy to use and doesn't have unnecessary features is a difficult job. It requires analysis of internal and external editor tools, prioritizing their importance to such a product. I have built Pie with these aspects in mind. Pie is a general-purpose code editor that provides a consolidated suite of the most used development features, accessible from an intuitive user interface. This tool allows programmers to focus on their code, instead of constantly clicking buttons, pressing hotkeys, or skimming through the editor's documentation. The entire feature analysis has been done by me, while using Pie on a daily basis as a complete replacement for other editors.

Pie, written entirely in C\#, has been built using Microsoft's Windows Forms ("Winforms") framework and has been designed to completely replace today's popular IDEs, such as Visual Studio Code or IntelliJ, whenever their features are just too much for a certain task. The solution has also been designed as a potential replacement for smaller-scale text editors, such as Notepad++, with various integrated text editing capabilities.

The demo shown in this paper will emphasize Pie's intuitive user interface along with its main features, highlighting some of the product's use cases. A comparison between Pie and state-of-the-art text and code editors will also be performed, focusing on what my product has to offer.
\vfill